Marlowe: financial contracts on blockchain

### 2018-05-18 tex/paper.tex
Made modes in Section 3 and 7 as proposed by PLS; removed 'other approaches' section from section 6

\section{Introduction}

@@

In this section, we revisit the escrow example that we discussed briefly in the introduction, and show how we can make 
Marlowe contracts that are easier to write, read, and understand, by embedding them into Haskell code, taking advantage 
of the fact that Marlowe contracts are implemented as Haskell terms. We used Haskell because it is the language in 
which Marlowe is implemented, but it would be easy to embed Marlowe in any other languages. It would only be necessary 
to translate its primitives into a data type in that language. In Meadow, we use Fay embedding, which is a subset of 
Haskell (see Section~\ref{section:tool}).

The example contract we use through this section implements an escrow contract. Let us imagine that participant 
\haskellinline{1} is buying some item from participant \haskellinline{2}. On one hand, if participant \haskellinline{1} 
pays before receiving the item, there is a risk that participant \haskellinline{2} will never give \haskellinline{1} the 
item. On the other hand, if participant \haskellinline{2} gives participant \haskellinline{1} the item before 
\haskellinline{1} pays,  there is a risk that \haskellinline{1} will never pay.

The escrow mechanism allows participant \haskellinline{1} to deposit the money into a contract, in a way that the money 
will only be released when two out of three participants agree on whether participant \haskellinline{2} has indeed given 
participant \haskellinline{1} the item. The third participant (participant \haskellinline{3}) is supposed to be a 
neutral third party that will decide in case of dispute.

This way, if participants \haskellinline{1} and \haskellinline{2} are honest, they will just agree on the result of the 
transaction and \haskellinline{3} will not need to do anything. If \haskellinline{1} and \haskellinline{2} disagree, 
participant \haskellinline{3} will be able to choose whether the money must go to \haskellinline{1} or to 
\haskellinline{2}.

In addition to this, in our implementation, the money paid for the item is 450 ADA, it must be committed by participant 
\haskellinline{1} before block \haskellinline{10}, and it will be refunded to participant \haskellinline{1} if there is 
no consensus before block \haskellinline{90}.

@@

Reflecting on the example, we can see that using Haskell definitions has made the contract substantially more 
comprehensible; it will also support more efficient operation, if we were to replace the repeated  expression 
\haskellinline{two_chose 1 2 3 1} above by a \haskellinline{where} clause, 
\begin{haskellcode}
escrow = ...        (When (OrObs (two_chose 1 2 3 0)
                                 chose_refund)
                          90
                          (Choice chose_refund ...
                           ...
         where chose_refund = two_chose 1 2 3 1                 
\end{haskellcode}
then the repeated computation of the expression will be avoided.
In the next section, we cover the online contract visualisation and interaction tool for Marlowe.

@@

The \ae{}ternity system~\cite{aeternity} provides multiple languages and VMs~\cite{Stenman-CODE-BEAM}: the functional language  Sophia (akin to 
Reason) and the Functional Typed Warded Virtual Machine (FTWVM) for safe ``system level programming'', the language 
Varna and the HLM for simple contracts, and a (port of) Solidity and the EVM for compatibility with Ethereum. In a 
similar way, Cardano provides support for IELE~\cite{IELE}, a rational reconstruction of the EVM, and thus for Solidity 
contracts too.

@@

\paragraph{Other approaches}

There are a variety of other approaches to scripting, including business process notations, such as the OMG's BPMN; 
process calculi such as Rholang; logic rules (in Formal Contract Logic), and finite state machines. More details of 
these are provided in~\cite{cryptoeprint:2016:1156}.

@@

We would like to thank Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles Hoskinson, Aggelos Kiayias, % Darryl McAdams, 
Bruno Woltzenlogel Paleo, Rebecca Valentine and Phil Wadler for their collaboration, and to colleagues from the Universities of 
Kent and Leicester for their comments on this work.

>>>>

\section{Introduction}
\label{sec:intro}

@@

In this section, we revisit the escrow example that we discussed briefly in the introduction, and show how we can make 
Marlowe contracts that are easier to write, read, and understand, by embedding them into Haskell code, taking advantage 
of the fact that Marlowe contracts are implemented as Haskell terms.

We used Haskell because it is the language in 
which Marlowe is implemented, but it would be easy to embed Marlowe in any other language. It would only be necessary 
to translate its primitives into a data type in that language. In Meadow we use Fay, which is a subset of 
Haskell (see Section~\ref{section:tool}).

The example we use through this section implements an escrow contract, as first introduced in Section~\ref{sec:intro}. 
%
%Let us imagine that participant 
%\haskellinline{1} is buying some item from participant \haskellinline{2}. On one hand, if participant \haskellinline{1} 
%pays before receiving the item, there is a risk that participant \haskellinline{2} will never give \haskellinline{1} the 
%item. On the other hand, if participant \haskellinline{2} gives participant \haskellinline{1} the item before 
%\haskellinline{1} pays,  there is a risk that \haskellinline{1} will never pay.
%
The escrow mechanism allows participant \haskellinline{1} to deposit the money into a contract, in a way that the money 
will only be released when two out of three participants agree on whether participant \haskellinline{2} has indeed given 
participant \haskellinline{1} the item.

The third participant (participant \haskellinline{3}) is supposed to be a 
neutral third party that will decide in case of dispute.
This way, if participants \haskellinline{1} and \haskellinline{2} are honest, they will just agree on the result of the 
transaction and \haskellinline{3} will not need to do anything. If \haskellinline{1} and \haskellinline{2} disagree, 
participant \haskellinline{3} will be able to choose whether the money must go to \haskellinline{1} or to 
\haskellinline{2}.

In our implementation we make things more specific: the money paid for the item is 450 ADA, and it must be committed by participant 
\haskellinline{1} before block \haskellinline{10}; it will be refunded to participant \haskellinline{1} if there is 
no consensus before block \haskellinline{90}.

@@

Reflecting on the example, we can see that using Haskell definitions has made the contract substantially more 
comprehensible. While our current implementation does not do this, it is possible to modify the embedding to support more efficient operation, by preserving \emph{sharing} in the host language. For example, if we were to replace the repeated  expression 
\haskellinline{two_chose 1 2 3 1} by a \haskellinline{where} clause, 
\begin{haskellcode}
escrow = ...        (When (OrObs (two_chose 1 2 3 0)
                                 chose_refund)
                          90
                          (Choice chose_refund ...
                           ...
         where chose_refund = two_chose 1 2 3 1                 
\end{haskellcode}
then the repeated computation of the expression could be avoided.
%In the next section, we cover the online contract visualisation and interaction tool for Marlowe.

@@

The \ae{}ternity system~\cite{aeternity} provides multiple languages and VMs~\cite{Stenman-CODE-BEAM}: the functional language  Sophia (akin to 
Reason) and the Functional Typed Warded Virtual Machine (FTWVM) for safe ``system level programming'', the language 
Varna and the HLM for simple contracts, and a (port of) Solidity and the EVM for compatibility with Ethereum. In a 
similar way, Cardano provides support for IELE~\cite{IELE}, a rational reconstruction of the EVM, and thus for Solidity too.

@@

%\paragraph{Other approaches}
%
%There are a variety of other approaches to scripting, including business process notations, such as the OMG's BPMN; 
%process calculi such as Rholang; logic rules (in Formal Contract Logic), and finite state machines. More details of 
%these are provided in~\cite{cryptoeprint:2016:1156}.

@@

We would like to thank IOHK for supporting us: not only has this led us to work on Marlowe, but we have benefitted hugely from collaboration with colleagues including Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles Hoskinson, Aggelos Kiayias, % Darryl McAdams, 
Bruno Woltzenlogel Paleo, Rebecca Valentine and Phil Wadler. We are  very grateful to Thomas Arts and colleagues from \ae{}ternity for their convivial discussions on blockchain and contracts in G\"oteborg, and to colleagues from the Universities of Kent and Leicester for their comments too.
